### 2018-09-11 outputs/docs/mentions/paper.Rmd
Start papers and get missing PDFs in HoR and start QLD

---
title: 'Slide into my mentions: Who gets incumbancy advantage in Australia?'
author: "Rohan Alexander"
date: "10/09/2018"
abstract: "I examine Australian hansards to create measures of a politician's importance within parliament. I then show these measures are positively correlated with incumbancy advantage."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Measuring the advantage of incumbency has been of interest to scholars since the second election. Bias makes it difficult to

Who gets kicked out at an election? Do people at the centre get kicked out? Measure those at the centre by those that have high mentions.