### 2018-10-02 scripts/to_monica/exploratory_2.Rmd
start of dirichlet modeling by MA

---
title: "Dirichlet?"
author: "Monica Alexander"
date: "10/1/2018"
output: html_document
---

Try and get this dirichlet thing going

```{r, message=F}
library(tidyverse)
library(lubridate)
library(boot)
library(rjags)
library(R2jags)
library(distortr)
```

Read in the data

```{r}
d <- read_csv("gammas_model_prevalence_years_k_20.csv")
dr <- d %>% 
  filter(document > ymd("2016-01-01"))
```

Plot these just to visualize. Note that the topics are really dominated by two particular ones.

```{r}
dr %>% 
  ggplot(aes(document, gamma, fill = factor(topic))) + geom_bar(stat = "identity")
```

Get info about sitting period, day number in sitting period etc. In this period we have 10 different sitting days. Sitting periods usually only have 8 days in them.

```{r}
dr_wide <- dr %>% 
  spread(topic, gamma)

dr_wide <- dr_wide %>% mutate(day_diff = document - lag(document))

k <- 1
dr_wide$group <- NA
for(i in 1:nrow(dr_wide)){
  if(is.na(dr_wide$day_diff[i])|dr_wide$day_diff[i]<7){
    k <- k
  }
  else{
    k <- k+1
  }
  dr_wide$group[i] <- k
}

dr_wide <- dr_wide %>% group_by(group) %>% 
  mutate(day_number = row_number())

```

Make back into long format. Now we have some covariates!

```{r}
dr_long <- dr_wide %>% 
  gather(topic, gamma, -document, -group, -day_number, -day_diff) %>% 
  arrange(document)
```

Think about a simple model here. The dependent variable is $\gamma_{stp}$, for sitting period $s$, day $t$ and topic $p$. This could be a function of: sitting period, the gamma of the same topic the day before, the year....

Start off with

$$
\gamma_{st,} \sim Dirichlet(\mu_{st,})\\
\log \mu_{stp} = \alpha_{sp} 
$$

So just linear in time for each topic and sitting period. Now need to get the data in the right form to fit:

```{r}
ntopics <- 20
y.stp <- array(NA, c(max(dr_long$group),max(dr_long$day_number), 20))

for (i in 1:ntopics){
  y.stp[,,i] <- dr_long %>% 
  ungroup() %>% 
  filter(topic==i) %>% 
  dplyr::select(-topic, -day_diff, -document) %>% 
  spread(day_number, gamma) %>% 
  dplyr::select(-group) %>% 
  as.matrix()
}

nsittings <- max(dr_long$group)
ndays.s <- dr_long %>% group_by(group) %>% summarise(days = max(day_number)) %>% dplyr::select(days) %>% pull()

```

Run the model:

```{r}
jags.data <- list(nsittings = nsittings,
                  ndays.s = ndays.s, 
                  ntopics = ntopics, 
                  y.stp = y.stp)
parnames <- c("alpha", "mu.stp")

mod <- jags(data = jags.data, 
            #n.iter = 5000,
            parameters.to.save=parnames,
            model.file = "model_dirichlet.txt")

max(mod$BUGSoutput$summary[,"Rhat"])
mcmc.array <- mod$BUGSoutput$sims.array
```

```{r}
df_res <- c()
for(i in 1:nsittings){
  for(j in 1:ndays.s[i]){
    for(k in 1:ntopics)
    df_res <- rbind(df_res, tibble(group = i, day = j, topic = k, 
                                   document =  dr_long$document[dr_long$group==i&dr_long$day_number==j&dr_long$topic==k],
                                   median = (median(mcmc.array[,,paste0("mu.stp[",i,",",j, ",",k,"]")])),
                                   upper = (quantile(mcmc.array[,,paste0("mu.stp[",i,",",j, ",",k,"]")], 0.975)),
                                   lower = (quantile(mcmc.array[,,paste0("mu.stp[",i,",",j, ",",k,"]")], 0.025))
    ))
  }
}
```

Plot:

```{r}
df_res %>% 
ggplot(aes(day, median, fill = factor(topic))) + geom_line(aes(color = factor(topic))) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  facet_wrap(~group)+
  geom_point(data = dr_long, aes((day_number), gamma, color = factor(topic)))
```