### 2018-10-04 scripts/from_monica/exploratory_4.Rmd
progress on the means model

---
title: "Dirichlet with government groups"
author: "Monica Alexander"
date: "10/3/2018"
output: html_document
---

Now we have a model going, let's try and add in some groupings based on changes in government. Hopefully then we will be able to visualize the 'average' topic distributions across governments. Let's try

$$
\gamma_{stp} \sim Dirichlet(\mu_{stp}) \\ 
\log \mu_{stp} = \alpha_{sp} + \delta_{stp}\\
\alpha_{s1} = 0 \\
\alpha_{sp} \sim N(\eta_{g[s]}, \sigma^2_g) \text { for } p>1 \\
\delta_{s,t,p} \sim N(\delta_{s,t-1,p}, \sigma^2_{\delta}) \\
\sigma^2_{\delta} \sim U(0, 40)
$$

```{r}
library(tidyverse)
library(lubridate)
library(rjags)
library(R2jags)
source("plottrace.R")
```

```{r}
d1 <- read_csv("dates_of_gov_and_election.csv")
d2 <- read_csv("gammas_model_prevalence_years_k_20.csv")
d <- d2 %>% 
  left_join(d1 %>% rename(document = allDates))
```

```{r}
dr <- d %>% 
  filter(electionCounter %in%41:42)
```

```{r}
dr %>% 
  ggplot(aes(document, gamma, fill = factor(topic))) + geom_bar(stat = "identity")
```

```{r}
unique(dr$electionCounter)
```

Get info about sitting period, day number in sitting period.

```{r}
dr_wide <- dr %>% 
  spread(topic, gamma)

dr_wide <- dr_wide %>% mutate(day_diff = document - lag(document))

k <- 1
dr_wide$group <- NA
for(i in 1:nrow(dr_wide)){
  if(is.na(dr_wide$day_diff[i])|dr_wide$day_diff[i]<7){
    k <- k
  }
  else{
    k <- k+1
  }
  dr_wide$group[i] <- k
}

dr_wide <- dr_wide %>% group_by(group) %>% 
  mutate(day_number = row_number())

dr_long <- dr_wide %>% 
  gather(topic, gamma, -document, -group, -day_number, -day_diff, -electionCounter, -governmentChangeDate) %>% 
  arrange(document)

```

Now need to get the data in the right form to fit:

```{r}
ntopics <- 20
y.stp <- array(NA, c(max(dr_long$group), max(dr_long$day_number), 20))

for (i in 1:ntopics){
  y.stp[,,i] <- dr_long %>% 
  ungroup() %>% 
  filter(topic==i) %>% 
  dplyr::select(-topic, -day_diff, -document, -electionCounter, -governmentChangeDate) %>% 
  spread(day_number, gamma) %>% 
  dplyr::select(-group) %>% 
  as.matrix()
}

nsittings <- max(dr_long$group)
ndays.s <- dr_long %>% group_by(group) %>% summarise(days = max(day_number)) %>% dplyr::select(days) %>% pull()

election.s <- (dr_long %>% group_by(group) %>% summarise(election = electionCounter[row_number()==1]) %>% dplyr::select(election) %>% pull()) - min(dr_long$electionCounter) + 1

nelections <- length(unique(dr_long$electionCounter))

```

Run the model:

```{r}
jags.data <- list(nsittings = nsittings,
                  nelections = nelections,
                  ndays.s = ndays.s, 
                  election.s = election.s,
                  ntopics = ntopics, 
                  y.stp = y.stp)
parnames <- c("alpha", "mu.sp", "mu.alpha")

mod <- jags(data = jags.data, 
            #n.iter = 5000,
            parameters.to.save=parnames,
            model.file = "model_dirichlet_means.txt")

max(mod$BUGSoutput$summary[,"Rhat"])
mcmc.array <- mod$BUGSoutput$sims.array
PlotTrace("mu.alpha[1,1]", mcmc.array)
```

Extract and plot the results:

```{r}
df_res <- c()
for(i in 1:nsittings){
  for(j in 1:ndays.s[i]){
    for(k in 1:ntopics)
    df_res <- rbind(df_res, tibble(group = i, day = j, topic = k, 
                                   document =  dr_long$document[dr_long$group==i&dr_long$day_number==j&dr_long$topic==k],
                                   median = (median(mcmc.array[,,paste0("mu.sp[",i,",",k,"]")])),
                                   upper = (quantile(mcmc.array[,,paste0("mu.sp[",i,",",k,"]")], 0.975)),
                                   lower = (quantile(mcmc.array[,,paste0("mu.sp[",i,",",k,"]")], 0.025))
    ))
  }
}
```

```{r}
df_res %>% 
ggplot(aes(day, median, fill = factor(topic))) + geom_line(aes(color = factor(topic))) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) + 
  facet_wrap(~group)+
  geom_point(data = dr_long, aes((day_number), gamma, color = factor(topic)))
```

Pull out the means by election and have a look

```{r}
mu_res <- c()
for(i in 1:nelections){
  for(j in 1:ntopics)
    mu_res <- rbind(mu_res, tibble(election = i, topic = j, 
                                   median = (median(mcmc.array[,,paste0("mu.alpha[",i,",",j,"]")])),
                                   upper = (quantile(mcmc.array[,,paste0("mu.alpha[",i,",",j,"]")], 0.975)),
                                   lower = (quantile(mcmc.array[,,paste0("mu.alpha[",i,",",j,"]")], 0.025))
    ))
}
```

```{r}
mu_res %>% 
  ggplot(aes(election, median, color = factor(election))) + 
  geom_point() + geom_errorbar(aes(ymin = lower, ymax = upper)) + 
  facet_wrap(~topic)
```

Have a look at the alphas

```{r}
alpha_res <- c()
for(i in 1:nsittings){
  for(j in 1:ntopics)
    alpha_res <- rbind(alpha_res, tibble(sitting = i, topic = j, election = election.s[i],
                                   median = (median(mcmc.array[,,paste0("alpha[",i,",",j,"]")])),
                                   upper = (quantile(mcmc.array[,,paste0("alpha[",i,",",j,"]")], 0.975)),
                                   lower = (quantile(mcmc.array[,,paste0("alpha[",i,",",j,"]")], 0.025))
    ))
}
```

```{r}
ggplot(alpha_res, aes(median, group = factor(election), color = factor(election))) + 
  geom_density() + facet_wrap(~topic)
```

```{r}
alpha_res %>% 
  ggplot(aes(sitting, median, color = factor(election))) + geom_point() + facet_wrap(~topic)
```